---
title: "H9_G1_KOR_pairwise_nodedistance"
author: "Jiani Chen"
date: "9/9/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

list.of.packages <- c("ggplot2","ggtree","ape","dplyr","tidyr")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)


#library(treeio)
library(ggplot2)
library(ggtree)
#library(phytools)
library(ape)
library(dplyr)
#library(tidytree)
#library('matrixStats')
library(tidyr)
#library(magrittr)
```

# G1 clades defined from annotation files

```{r, warning=FALSE,echo=FALSE}
G1<-read.tree(file="01_data/G1conOG.fas.treefile")

p<-ggtree(G1)
meta_G1<- read.table("01_data//G1_clade.txt", sep="\t", header=TRUE,check.names=FALSE, stringsAsFactor=F, fill=TRUE)

p%<+% meta_G1 +
  geom_tippoint(aes(color=clade), size=2, alpha=.75) +
  geom_treescale(x=0.03, y=500,linesize= 1,,width=0.01)+
##  geom_text2(aes(label=label),color="blue",size =3)+ ## this is used to list the node number
  geom_text2(aes(label=label, subset=(node %in% c(1414,1828,2036,2084,2371,2389,2397,2527,2568,2694))),color="blue",size =3)+
  geom_point2(aes(label=label, subset=(node %in% c(1414,1828,2036,2084,2371,2389,2397,2527,2568,2694))),color="blue",size =2)+
  scale_color_brewer(palette = "Set3",name="G1 annotation") 
```

# Calulate the maximal pair-wise node distance within the assigned clades

```{r, echo=FALSE, warning=FALSE,message=FALSE}
max_inclade<-function(tree,node){
  clade<-extract.clade(tree,node)
  clade_nodedis<-dist.nodes(clade)%>%
    as.data.frame()
  
  clade_nodedis$node1 <-rownames(clade_nodedis)
  
  clade_nodedis2<-pivot_longer(clade_nodedis,cols=1:nrow(clade_nodedis),names_to="node2",values_to ="node_dist")
  parentlist<-unique(as_tibble(clade)$parent)
  as_tibble(clade)%>%
    filter(as.numeric(label) >=70)->supp_node
  nodeList<-supp_node$node  
  clade_nodedis3 <-clade_nodedis2 %>% 
    subset(node1 %in% nodeList) %>% subset(node2 %in% nodeList) %>%
    subset(node1 %in% parentlist) %>% subset(node2 %in% parentlist)
  return(max(clade_nodedis3$node_dist))
  
}


data.frame("clade"=c("G1-A","G1-B","G1-C","G1-D","G1-E","G1-F","G1-G","G1-H","G1-I","G1-L"),
           "max within-clade pair-nodedist"=c(max_inclade(G1,1414),max_inclade(G1,1828),
                                              max_inclade(G1,2036),max_inclade(G1,2083),
                                              max_inclade(G1,2371),max_inclade(G1,2389),
                                              max_inclade(G1,2397),max_inclade(G1,2527),
                                              max_inclade(G1,2568),max_inclade(G1,2694)))

```

# Simulation for G1-clade

```{r, echo=FALSE,results='hide', warning=FALSE,message=FALSE}
## test to perform genotype assignment with node distance
## simulation different cutoff value , number of cluster

G1_nodedis<-dist.nodes(G1)%>%
  as.data.frame()

G1_nodedis$node1 <-rownames(G1_nodedis)
G1_nodedis2<-pivot_longer(G1_nodedis,cols=1:2792,names_to="node2",values_to ="node_dist")

as_tibble(G1)%>%
  filter(as.numeric(label) >=70)->supp_node

parentlist<-unique(as_tibble(G1)$parent)
nodeList<-supp_node$node    
G1_nodedis3 <-G1_nodedis2 %>% 
  subset(node1 %in% nodeList) %>% subset(node2 %in% nodeList) %>%
  subset(node1 %in% parentlist) %>% subset(node2 %in% parentlist) 

G1_nodedis4 <- pivot_wider(G1_nodedis3, names_from = node2, values_from = node_dist)  

rownames(G1_nodedis4)<-G1_nodedis4$node1
G1_nodedis5<-G1_nodedis4 %>% 
  select(-node1)%>%
  as.dist()

G1_simout<-data.frame(cutoff=double(),
                       ngroup=integer())

for (i in seq(0.001,0.2,by=0.001)){
  hr <- hclust(G1_nodedis5, method = "complete", members=NULL)
  myhcl <- cutree(hr, h=i)
  out<-c(i,length(unique(myhcl)))
  G1_simout <-rbind(G1_simout,out)
  
}

ggplot(G1_simout,aes(x=G1_simout$X0.001,y=G1_simout$X807))+
  geom_point()+
  geom_smooth()+
  theme_bw()+
  xlab("Cut-off vlaue of max within clade pair-wise node distance")+
  ylab("Number of clades to be assign")
```

## I re-assign the clades with maximal pair-wise node distance within clade as 0.16 subs/site
## Similar resultes from the clades defined in annotation file
### G1-D clade, small propotion of sequences was unable to classified with this approach ( type 8, green) 


```{r, echo=FALSE,warning=FALSE,results='hide'}

map2color<-function(x,pal,limits=NULL){
  if(is.null(limits)) limits=range(x)
  pal[findInterval(x,seq(limits[1],limits[2],length.out=length(pal)+1), all.inside=TRUE)]
} 
hr <- hclust(G1_nodedis5, method = "complete", members=NULL)
myhcl <- cutree(hr, h=0.16) ######### edit the cut-off 

G1_sum<-data.frame()
G1_nodename<-G1_nodedis4$node1 
G1_sum<-cbind(myhcl)%>%as.data.frame()%>%
  mutate(node=as.integer(G1_nodename))%>%
  mutate(color=map2color(myhcl,rainbow(200)))

nnode<-c()
for(i in as.numeric(G1_sum$node)){
  temp<-extract.clade(G1,i)
  nnode<-c(nnode,temp$Nnode)
  
}

G1_sum<-G1_sum%>%
  mutate(nnode=nnode)%>%
  group_by(myhcl) %>% 
  filter(nnode == max(nnode)) 

## get the tips from node
G1_tip<-data.frame()
G1_temp<-extract.clade(G1,1400)
G1_temp2<-extract.clade(G1,1403)
G1_temp3<-extract.clade(G1,1408)
G1_temp4<-extract.clade(G1,1414)
#G1_temp5<-extract.clade(G1,1812)
G1_temp6<-extract.clade(G1,1829)
G1_temp7<-extract.clade(G1,2036)
G1_temp8<-extract.clade(G1,2086)
G1_temp9<-extract.clade(G1,2372)
G1_temp10<-extract.clade(G1,2390)
G1_temp11<-extract.clade(G1,2399)
G1_temp12<-extract.clade(G1,2694)

G1_tip<-cbind(G1_temp$tip.label)%>%
  as.data.frame()%>%
  mutate(genotype="type1")%>%
  mutate(genotype=ifelse(V1 %in% G1_temp2$tip.label,"type2",genotype)) %>%
  mutate(genotype=ifelse(V1 %in% G1_temp3$tip.label,"type3",genotype)) %>%
  mutate(genotype=ifelse(V1 %in% G1_temp4$tip.label,"type4",genotype)) %>%
  #mutate(genotype=ifelse(V1 %in% G1_temp5$tip.label,"type5",genotype))%>%
  mutate(genotype=ifelse(V1 %in% G1_temp6$tip.label,"type5",genotype))%>%
  mutate(genotype=ifelse(V1 %in% G1_temp7$tip.label,"type6",genotype))%>%
  mutate(genotype=ifelse(V1 %in% G1_temp8$tip.label,"type7",genotype))%>%
  mutate(genotype=ifelse(V1 %in% G1_temp9$tip.label,"type8",genotype))%>%
  mutate(genotype=ifelse(V1 %in% G1_temp10$tip.label,"type9",genotype))%>%
  mutate(genotype=ifelse(V1 %in% G1_temp11$tip.label,"type10",genotype))%>%
  mutate(genotype=ifelse(V1 %in% G1_temp12$tip.label,"type11",genotype))
colnames(G1_tip)<-c("taxa","genotype")


p%<+% G1_tip +
  geom_tippoint(aes(color=genotype), size=2, alpha=.75) +
  geom_treescale(x=0.03, y=500,linesize= 1,,width=0.01)+
  scale_color_brewer(palette = "Set3",name="G1 annotation with pair-wise node distance") 

```

# Analysis for KOR clade
# Kor clades defined from annotation files

```{r, warning=FALSE,echo=FALSE}
kor<-read.tree(file="01_data/tree-Korean-topology-with-outgroup-selection.nwk")

p<-ggtree(kor)
meta_kor<- read.table("01_data/Korean-Groups.txt", sep="\t", header=TRUE,check.names=FALSE, stringsAsFactor=F, fill=TRUE)

p%<+% meta_kor +
  geom_tippoint(aes(color=Group), size=2, alpha=.75) +
  geom_treescale(x=0.03, y=500,linesize= 1,,width=0.01)+
  geom_text2(aes(label=label, subset=(node %in% c(585,773,887,911,915,980,1097,1110))),color="blue",size =3)+
  geom_point2(aes(label=label, subset=(node %in% c(585,773,887,911,915,980,1097,1110))),color="blue",size =2)+
  scale_color_brewer(palette = "Set3",name="Kor annotation") 
```

# Calulate the maximal pair-wise node distance within the assigned clades

```{r, echo=FALSE, warning=FALSE,message=FALSE}
max_inclade<-function(tree,node){
  clade<-extract.clade(tree,node)
  clade_nodedis<-dist.nodes(clade)%>%
    as.data.frame()
  
  clade_nodedis$node1 <-rownames(clade_nodedis)
  
  clade_nodedis2<-pivot_longer(clade_nodedis,cols=1:nrow(clade_nodedis),names_to="node2",values_to ="node_dist")
  parentlist<-unique(as_tibble(clade)$parent)
  #as_tibble(clade)%>%
   # filter(as.numeric(label) >=70)->supp_node
  nodeList<-supp_node$node  
  clade_nodedis3 <-clade_nodedis2 %>% 
    #subset(node1 %in% nodeList) %>% subset(node2 %in% nodeList) %>%
    subset(node1 %in% parentlist) %>% subset(node2 %in% parentlist)
  return(max(clade_nodedis3$node_dist))
  
}

data.frame("clade"=c("KOR-A","KOR-B","KOR-C","KOR-D","KOR-E","KOR-F","KOR-G","KOR-H"),
            "max within-clade pair-nodedist"=c(max_inclade(kor,585),max_inclade(kor, 773),
                                               max_inclade(kor,887),max_inclade(kor, 911),
                                               max_inclade(kor,915),max_inclade(kor,980),
                                               max_inclade(kor,1097),max_inclade(kor,1110)))
```

## simulation with different maximal within clade pair-wise node distances 

```{r, echo=FALSE,results='hide', warning=FALSE,message=FALSE}

kor_nodedis<-dist.nodes(kor)%>%
  as.data.frame()

kor_nodedis$node1 <-rownames(kor_nodedis)
kor_nodedis2<-pivot_longer(kor_nodedis,cols=1:1149,names_to="node2",values_to ="node_dist")



parentlist<-unique(as_tibble(kor)$parent)
nodeList<-supp_node$node    
kor_nodedis3 <-kor_nodedis2 %>% 
  #subset(node1 %in% nodeList) %>% subset(node2 %in% nodeList) %>%
  subset(node1 %in% parentlist) %>% subset(node2 %in% parentlist) 

kor_nodedis4 <- pivot_wider(kor_nodedis3, names_from = node2, values_from = node_dist)  

rownames(kor_nodedis4)<-kor_nodedis4$node1
kor_nodedis5<-kor_nodedis4 %>% 
  select(-node1)%>%
  as.dist()

kor_simout<-data.frame(cutoff=double(),
                       ngroup=integer())

for (i in seq(0.001,0.2,by=0.001)){
  hr <- hclust(kor_nodedis5, method = "complete", members=NULL)
  myhcl <- cutree(hr, h=i)
  out<-c(i,length(unique(myhcl)))
  kor_simout <-rbind(kor_simout,out)
  
}

ggplot(kor_simout,aes(x=kor_simout$X0.001,y=kor_simout$X366))+
  geom_point()+
  geom_smooth()+
  theme_bw()+
  xlab("Cut-off vlaue of max within clade pair-wise node distance")+
  ylab("Number of clades to be assign")
```


## I re-assign the clades with maximal pair-wise node distance within clade as 0.16 subs/site
## Similar resultes from the clades defined in annotation files, only differences in KOR-A
### KOR-A was divided into two propotion with pair-wise node distance criteria (type1 and type 3)


```{r, warning=FALSE,echo=FALSE}
hr <- hclust(kor_nodedis5, method = "complete", members=NULL)
myhcl <- cutree(hr, h=0.16)

kor_sum<-data.frame()
kor_nodename<-kor_nodedis4$node1 
kor_sum<-cbind(myhcl)%>%as.data.frame()%>%
  mutate(node=as.integer(kor_nodename))%>%
  mutate(color=map2color(myhcl,rainbow(200)))

nnode<-c()
for(i in as.numeric(kor_sum$node)){
  temp<-extract.clade(kor,i)
  nnode<-c(nnode,temp$Nnode)
  
}

kor_sum<-kor_sum%>%
  mutate(nnode=nnode)%>%
  group_by(myhcl) %>% 
  filter(nnode == max(nnode)) 

## get the tips from node
kor_tip<-data.frame()
kor_temp<-extract.clade(kor,576)
kor_temp2<-extract.clade(kor,582)
kor_temp3<-extract.clade(kor,585)
kor_temp4<-extract.clade(kor,592)
kor_temp5<-extract.clade(kor,911)
kor_temp6<-extract.clade(kor,915)
kor_temp7<-extract.clade(kor,980)
kor_temp8<-extract.clade(kor,1097)
kor_temp9<-extract.clade(kor,1110)


kor_tip<-cbind(kor_temp$tip.label)%>%
  as.data.frame()%>%
  mutate(genotype="type1")%>%
  mutate(genotype=ifelse(V1 %in% kor_temp2$tip.label,"type2",genotype)) %>%
  mutate(genotype=ifelse(V1 %in% kor_temp3$tip.label,"type3",genotype)) %>%
  mutate(genotype=ifelse(V1 %in% kor_temp4$tip.label,"type4",genotype)) %>%
  mutate(genotype=ifelse(V1 %in% kor_temp5$tip.label,"type5",genotype))%>%
  mutate(genotype=ifelse(V1 %in% kor_temp6$tip.label,"type6",genotype))%>%
  mutate(genotype=ifelse(V1 %in% kor_temp7$tip.label,"type7",genotype))%>%
  mutate(genotype=ifelse(V1 %in% kor_temp8$tip.label,"type8",genotype))%>%
  mutate(genotype=ifelse(V1 %in% kor_temp9$tip.label,"type9",genotype))

colnames(kor_tip)<-c("taxa","genotype")

p%<+% kor_tip +
  geom_tippoint(aes(color=genotype), size=2, alpha=.75) +
  geom_treescale(x=0.03, y=500,linesize= 1,,width=0.01)+
  scale_color_brewer(palette = "Set3",name="kor annotation with pair-wise node distance") 
```

